---
title: "2019 Field Work Summary"
output: pdf_document
---

```{r loadDat, include=FALSE}
# adjust as desired:
knitr::opts_chunk$set(
  fig.asp = 0.618,
  fig.width = 9,
  out.width = "6in",
  fig.path='Figs/', 
  include=FALSE,
  echo=FALSE, warning=FALSE, message=FALSE,
  dpi = 180,
  fig.align = "center",
  fig.pos = "htb"
)
options(kableExtra.latex.load_packages = FALSE)

library(tidyverse)
library(ggplot2)

setDat <- read.csv(here::here("data", "taggingData", "cleanSetData.csv")) 
chinDat <- read.csv(here::here("data", "taggingData", "cleanTagData.csv"))
```

The primary objective of our 2019 field season was to deploy acoustic telemetry tags that could be used to quantify stock-specific migration rates, depth distribution, and survival of adult Chinook salmon within resident killer whale critical habitat. Acoustics tags will be detected by passive receivers including a sparse grid array deployed along the continental shelf from the Columbia River to the Canadian border (managed by NOAA), a curtain array across Juan de Fuca Strait (OTN), and multiple smaller arrays throughout the Discovery Islands and Fraser River (UBC). Additionally, active receivers were deployed opportunistically by DFO personnel during killer whale surveys. Data from receivers are beginning to be downloaded by collaborators and should be available by late fall. 

The secondary objective of the 2019 field season was to estimate stock-specific abundance of adult Chinook salmon within killer whale critical habitat using catch per unit effort (CPUE) indices. This objective was prioritized below tagging because there were insufficient resources to develop a fully stratified random survey design that could fully account for variation among migrating stocks in spatial and temporal distribution.

We sampled adult Chinook salmon during five multi-day fishing events (July 11-August 24) using a 37-foot commercial troller. Fishing effort was concentrated offshore of Ucluelet at the northern border of resident killer whale critical habitat on the southwest coast of Vancouver Island (Figure 1). All Chinook salmon over 45 cm were landed, measured, sampled for stock identification, and fitted with a PIT tag (for detection in freshwater arrays in the Cowichan and Columbia rivers). A subset of Chinook salmon were tagged with acoustic transmitters with depth sensors that will be detected by the arrays described above. Non-target species (predominantly coho salmon and lingcod) were enumerated. All fish were released immediately following capture. 

```{r setMap, include=TRUE, fig.cap="Study area with trolling locations (points) colored by fishing date. The red polygon represents the newly designated resident killer whale habitat including La Perouse and Swiftsure banks."}
setMap <- readRDS(here::here("generatedData", "setMap.RDS"))
setMap
```

```{r summDat}
landed <- nrow(chinDat)
tagged <- nrow(chinDat %>% 
  filter(!is.na(acoustic)))
```

We landed and biosampled `r landed` maturing or mature Chinook salmon and deployed `r tagged` acoustic tags. Biosampled fish represented a relatively large size range and included both clipped and unclipped individuals. On average, unclipped fish were larger than clipped, although these differences were not significant (Figure 2).

```{r bioDat, include=TRUE, fig.cap="Size distributions of adipose clipped and unclipped Chinook salmon. Dots represent means and whiskers one standard deviation."}
dum <- chinDat %>% 
  filter(fl > 40)

stat_box_data <- function(y, upper_limit = max(chinDat$fl) * 1.25) {
  return( 
    data.frame(
      y = 0.95 * upper_limit,
      label = paste('count =', length(y), '\n',
                    'mean =', round(mean(y), 1), '\n')
    )
  )
}

ggplot(dum, aes(x = clip, y = fl, fill = clip, alpha = 0.1)) +
  geom_violin(trim = FALSE) +
  stat_summary(fun.data=mean_sdl, geom="pointrange") +
  stat_summary(fun.data = stat_box_data, geom = "text", hjust = 0.5, 
               vjust = 0.9) +
  scale_fill_viridis_d() +
  labs(x = "Adipose Clipped", y = "Fork Length") +
  samSim::theme_sleekX() +
  theme(legend.position = "none")
```

We are still awaiting genetic stock identification results and robust estimates of abundance will require more advanced modeling that accounts for sampling effort. However these preliminary data can still be used to visualize differences in spatial distribution among size classes or clipped (i.e. known hatchery origin) vs. unclipped individuals using raw catch data.

For example, initial catch data suggest a greater proportion of large individuals (>75 cm) are found offshore, while sublegal (presumably immature) fish are distributed further south and east (Figure 3). Conversely the spatial distributions of clipped and unclipped fish were similar.

```{r heatMap2, include=TRUE, fig.cap="Relative catch densities of different size classes of Chinook salmon. Points represent individual sets. Note that distributions based on raw data will be biased by sampling effort."}
chinHeatSize <- readRDS(here::here("generatedData", "heatMapSize.RDS"))
chinHeatSize
```

```{r heatMap3, include=TRUE, fig.cap="Relative catch densities of adipose clipped and unclipped Chinook salmon. Points represent individual sets. Note that distributions based on raw data will be biased by sampling effort."}
chinHeatClip <- readRDS(here::here("generatedData", "heatMapClip.RDS"))
chinHeatClip
```

Next steps include pairing the stock identified CPUE data we have collected with historic samples (2007-2015) collected from the west coast commercial troll fishery by South Coast. Combining these datasets will allow us to estimate stock-specific temporal and spatial distributions that are relatively robust to non-random sampling. The data generated by the acoustic tagging program will increase the resolution of these estimates. Furthermore we hope to partner with NOAA collaborators to extend the acoustic arrays to provide coverage on Swiftsure Bank, improving our estimates of spatial distribution and survival.

 